---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# geolink

<!-- badges: start -->
<!-- badges: end -->

Quickly identify areal identifiers like NUTS, ISO-3, GAUL, INSPIRE, AGS or FIPS and link them to their respective polygon geometries. Convenience wrapper around common API wrappers and standards to facilitate spatial analysis for people who don't know spatial analysis.


## Installation

You can install the development version of `{geolink}` from GitHub with:

``` r
# install.packages("pak")
pak::pak("jslth/geolink")
```

## Example

For basic use, `{geolink}` has a single function. Given a dataset with spatial identifiers, it automatically links the respective geometries:

```{r example}
library(geolink)
library(gt)

countrypops <- gt::countrypops
countrypops <- countrypops[countrypops$year == 2020, ]
countrypops <- enrich(countrypops, id_col = "country_code_2")
countrypops
```

```{r cars}
library(ggplot2)

ggplot(t2) +
    geom_sf(aes(fill = population)) +
    scale_fill_viridis_b("Population", transform = "log10") +
    theme_void()
```

If a country code is detected, `enrich()` will always convert it to ISO-3. However, the function can also handle other regional identifiers, e.g., FIPS codes:

```{r}
west_coast <- c("06", "41", "53")
west_coast <- enrich(west_coast, linker = "fips", crs = 4326)
plot(west_coast)
```

A number of "geolinkers" are pre-registered, but you can also register your own workflows:

```{r}
check <- function(ids) {
  grepl("x: -?[0-9]{0,3}, y: -?[0-9]{0,2}", ids)
}

parse <- function(ids) {
  strcapture(
    "x: (-?[0-9]{0,3}), y: (-?[0-9]{0,2})",
    ids,
    proto = list(x = double(), y = double())
  )
}

link <- function(.data, id_col, x, y, link_crs) {
  geom <- Map(\(x, y) sf::st_point(c(x, y)), x, y)
  geom <- sf::st_as_sfc(geom, crs = link_crs)
  .data$geometry <- geom
  sf::st_as_sf(.data)
}

register_geolinker(
  "coords_fmt",
  link = link,
  check = check,
  parse = parse
)
```

```{r}
test_data <- data.frame(id = c("x: 8, y: 47", "x: 9, y: 46", "x: 10, y: 45"))
enrich(test_data, link_crs = 4326)
```

